---
title: "Quick Intro to RMD: HUDM 5059"
author: "Rachel Lee"
date: "2023-09-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This file is to help you get started with [R Markdown](http://www.rstudio.org/docs/authoring/using_markdown) and [knitr](http://yihui.name/knitr/) in Rstudio.



## Getting started
The `#` symbol allows you to make titles. Explore using one, two, three `#` symbols to adjust the size.  

# Title 1
## Title 2
### Title 3

To skip a line, include a double space at the end of the sentence.  
To break a line, add `<br>` commands.
<br>
<br>
<br>

To work with R Markdown, if necessary:

* Install [R](http://www.r-project.org/)
* Install the lastest version of [RStudio](http://rstudio.org/download/)
* Install the latest version of the `knitr` package: `install.packages("knitr")`


## Prepare for analyses
```{r }
set.seed(1234)
library(ggplot2)
library(lattice)
```


## Basic console output
To insert an R code chunk, you can type it manually or just press `Chunks - Insert chunks` or use the shortcut key. This will produce the following code chunk:

    ```{r}
    
    ```

Pressing tab when inside the braces will bring up code chunk options.


The code chunk input and output is then displayed as follows:

```{r}
x <- 1:10
y <- round(rnorm(10, x, 1), 2)
df <- data.frame(x, y)
df
```

## Plots
Images generated by `knitr` are saved in a figures folder.

### Simple plot
Here is a basic plot using base graphics:

```{r}
plot(x)
```


### Multiple plots
You can include multiple plots in one code chunk:

```{r}
boxplot(1:10~rep(1:2,5))
plot(x, y)
```

### `ggplot2` plot
ggplot2 plots work well:

```{r}
qplot(x, y, data=df)
```

### `lattice` plot
As do lattice plots:

```{r}
xyplot(y~x)
```


## R Code chunk features
### Create Markdown code from R
The following code hides the command input (i.e., `echo=FALSE`), and outputs the content directly as code.
```{r}
cor(x,y)
```
Note that `echo=TRUE` by default. This shows your code AND the result.

```{r, echo=FALSE}
cor(x,y)
```
If you set `echo=FALSE`, then it simply shows your result. Please **DO NOT MANUTALLLY CHANGE THE SETTING BY TYPING IN `echo=FALSE` FOR OUR CLASS ASSIGNMENTS!!** We need to see your code!


### Control figure size
The following is an example of a smaller figure using `fig.width` and `fig.height` options.

```{r, fig.width=3, fig.height=3}
plot(x)
```

## Basic markdown functionality
For those not familiar with standard [Markdown](http://daringfireball.net/projects/markdown/), the following may be useful.
See the source code for how to produce such points.

### Dot Points
Simple dot points:

* Point 1
* Point 2
* Point 3

and numeric dot points:

1. Number 1
2. Number 2
3. Number 3

and nested dot points:

* A
    * A.1
    * A.2
* B
    * B.1
    * B.2


### Equations
Equations are included by using LaTeX notation and including them either between single dollar signs (inline equations) or double dollar signs (displayed equations).  

There are inline equations such as $y_i = \alpha + \beta x_i + e_i$.

And displayed formulas:

$$\frac{1}{1+\exp(-x)}$$


### Tables
Tables can be included using the following notation

A   |   B    | C
--- |  ---   | ---
1   |  Male  | Blue
2   | Female | Pink


There are other packages you can check out for embedding tables in RStudio, such as `kable`.
<br>
<br>
If you have any questions, please come to my office hours on Mondays 11-Noon.